Spin canonical divisors of log stacky curves

### 2015-06-23 stacky_write_up/spin_cring.tex
added genus 1 base cases loosely

\begin{prop}
\label{prop:sat_two_induction}
Let $(\sx, \Delta, L)$ and $(\sx', \Delta, L')$ be tame separably rooted log stacky spin curves. 
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted log stacky curves so that $\sx$ has signature 
$\sigma = (g;e_1,\ldots, e_r;\delta)$ with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.
Let $R_{L'} = R' = k[x]/I'= k[x_2, x_3
, x_5,\ldots, x_m]/I'$ and let $L = L' + \frac{1}{
3}P.$ Suppose $L$ and $L'$ satisfy $sat(Eff(L')) = sat(Eff(L)) = 2.$
Suppose $\deg x_i = i$ for $i \in \{2, 3\}$ and that the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_4 \prec 
		y_3$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{align}
\label{eqn:sat_two_add_generator}
	\begin{split}
		&\{ y_3^ax_2^b x_3^\epsilon \mid a \geq 0, b 
		\geq 0, \epsilon \in \{0, 1\}\} \\
		&\cup \{ y_3^ay_4 \mid a \geq 0 \},
	\end{split}
\end{align}

@@

\begin{align*}
	y_4x_j &- A_j y_3 w_j & \text{ so that } 2 \leq j \leq m \text{ and } \deg w_j = \deg x_j + 1\\
	y_4^2 &+ B y_3^2 x_2 \\
\end{align*}

@@

@@

\begin{lem}\label{lem:adding_points_saturation_2_to_1_with_log}
Let $(\sx, \Delta, \halfcan)$ and $(\sx, \Delta', L')$ be tame separably rooted log stacky spin 
curves with $0\ne \Delta=\Delta'+2P$ for some non-stacky point $P$ not in the support of $
\halfcan'$, meaning that $L=L'+P$.  Let $(g; e_1, \ldots, e_\tau; \delta+2)$ and $(g; e_1, 
\ldots, e_\tau, \delta)$ with $g\ge 2$ be their respective signatures.  Suppose 
$sat(Eff(\halfcan))=1$ and $sat(Eff(\halfcan'))=2$.  Further suppose $R_{\halfcan'} = k[x_2, 
x_3, \ldots, x_m]/I'$ with $\deg(x_2) = 2$ and $\deg(x_3) = 3$.  Then
\begin{enumerate}
\item[(a)] Then generic elements $y_1$ and $y_2$ in $(R_{\halfcan})_1$ and $
(R_{\halfcan})_2$ respectively, such that $-\ord_P(y_2)>0$ and $\{y_1^2, y_2\}$ is linearly 
independent, generate $R_\halfcan$ over $R_{\halfcan'}$.
\item[(b)] Equip $k[y_1,y_2]$ with $grlex$ so that $y_2 \prec y_1$ and equip $k[y_1, y_2, 
x_2, \ldots, x_m]$ with the block order so that $R_\halfcan=k[y_1, y_2, x_2, \ldots, x_m]/I$.  
Then
\begin{align*}
			in_\prec(I) &= in_\prec(I')k[x_2, \ldots, x_m, y_1, y_2] \\
			&+\langle y_2 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_2^2\rangle.
		\end{align*}
\item[(c)] Any set of minimal generators for $I'$ together with any set of relations with 
leading terms as above minimally generate $I$.
\end{enumerate}

\end{lem}
\begin{proof}
Since adding $P$ to $L'$, reduces saturation from 2 to 1, $h^0(X, \lfloor L \rfloor)=1$.  
Choose any $y_1\in H^0(X,\lfloor L\rfloor)$; we know by saturation arguments that $-
\ord_P(y_1)=1$.  Furthermore, since $\delta>0$, we use Riemann-Roch to deduce that for 
all $d>1$ $h^0(X,\lfloor dL\rfloor)=h^0(dL')+d$, so $(R_\halfcan)_d$ is a $d$ dimension $k
$-vector space over $(R_{\halfcan'})_d$.  In particular, this means we can generically 
choose an element $y_2\in H^0(X,\lfloor 2L\rfloor)$ linearly independent from $y_1^2$.  By 
dimension counting, we see that 
\[
	\langle y_1^a x_2^b x_3^c: a\ge 0, b\ge 0, c\in \{0,1\}\rangle
\]
\[
	\langle y_1^a y_2: a\ge 0\rangle
\]
form a $k$-basis for $R_\halfcan$ over $R_{\halfcan'}$.  This completes part (a).

To show part (b), we note that the monomials in the basis above generate $k[x,y]/\initial_
\prec(I)$.  Thus, it is sufficient to show that $J\subseteq \initial_\prec(I)$.  For $j\ge 2$, we 
can choose $w_j$ in degree $\deg(x_j)+1$ and $A_j\in k$ such that
\[
	y_2x_j-A_jy_3w_j
\]
has no pole at $P$ at thus lies in $R_{\halfcan'}$.  Similarly, we can choose $B$ and $C$ 
such that
\[y^2_4+By_1^2x_2+Cy_1x_3\]
has no pole at $P$, so it lies in $R_{\halfcan'}$.  These give us relations with initial terms 
as desired, concluding part (b).

Finally, (c) follows immediately from Lemma \ref{lem:minimal_quadratic}.

\end{proof}

\begin{lem}\label{lem:adding_points_saturation_2_to_1_no_log}
Let $(\sx, 2P, \halfcan)$ and $(\sx, 0, L')$ be tame separably rooted log stacky spin curves 
with $P$ not in the support of $\halfcan'$, meaning that $L=L'+P$.  Let $(g; e_1, \ldots, e_
\tau; 2)$ and $(g; e_1, \ldots, e_\tau, 0)$ with $g\ge 2$ be their respective signatures.  
Suppose $sat(Eff(\halfcan))=1$ and $sat(Eff(\halfcan'))=2$.  Further suppose 
$R_{\halfcan'} = k[x_2, x_3, \ldots, x_m]/I'$ with $\deg(x_2) = 2$ and $\deg(x_3) = 3$.  
Then
\begin{enumerate}
\item[(a)] Then generic elements $y_1$ and $y_3$ in $(R_{\halfcan})_1$ and $
(R_{\halfcan})_3$ respectively, such that $-\ord_P(y_3)>1$ and $\{y_1^3, y_2\}$ is linearly 
independent, generate $R_\halfcan$ over $R_{\halfcan'}$.
\item[(b)] Equip $k[y_1,y_3]$ with $grlex$ so that $y_3 \prec y_1$ and equip $k[y_1, y_3, 
x_2, \ldots, x_m]$ with the block order so that $R_\halfcan=k[y_1, y_3, x_2, \ldots, x_m]/I$.  
Then
\begin{align*}
			in_\prec(I) &= in_\prec(I')k[x_2, \ldots, x_m, y_1, y_2] \\
			&+\langle y_3 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_3^2
		\end{align*}
\item[(c)] Any set of minimal generators for $I'$ together with any set of relations with 
leading terms as above minimally generate $I$.
\end{enumerate}
\end{lem}
\begin{proof}
The proof follows similarly to that of lemma 
\ref{lem:adding_points_saturation_2_to_1_with_log}, with the change that $(R_
\halfcan)_2$ is only a degree 1 $k$-vector space over $(R_{\halfcan'})_2$ forcing a 
generator in degree 3 rather than degree 2.  
\end{proof}

@@

\begin{lem}
\label{lem:semicanonical_generation}\todo{include that this was as shown in Neves, with degree 6 omitted}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K$. Then, for $n \geq 5$, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective.
In particular, $R_L$ is generated in degree at most $6$.
\end{lem}
\begin{proof}
Therefore, to show there are no new generators in degree at least $7$, it suffices to show that for $n \geq 5$, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, let $s_1,s_2 \in H^0(2L) \cong H^0(K)$ be two independent sections. 
Since $K$ is basepoint free, so is the vector subspace $V = ks_1 \oplus k s_2 \subset H^0(K)$. 
Then, by the basepoint free pencil trick, we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$
We wish to show $f$ is surjective.
Note that $\dim \ker f = \dim H^0((n-2)L) = (n-3)(g-1)$ using Riemann--Roch and the assumption that $n \geq 5$.
Additionally, $\dim V \otimes H^0(nL) = 2 \cdot (n-1)(g-1)$, again using Riemann--Roch.
Therefore, $$\dim \im f = 2 \cdot (n-1)(g-1) -(n-3)(g-1) = (n+1)(g-1) = \dim H^0((n+2)L),$$
again using Riemann-Roch, implying $f$ is surjective.
\end{proof}

@@

\begin{cor}\label{cor:genus_0_posiive_delta}
Let $\halfcan$ be a stacky log spin canonical divisor of $(X,\Delta)$ with signature $(g;e_1,\ldots, e_n;\delta)$ such that $2\halfcan=K_\sx+\Delta$ with $\delta=\deg(\Delta)>0$.  Then $\halfcan$ is effective so we can inductively apply lemma \ref{lem:deg1_sat_ind}.
\end{cor}
\begin{proof}
Since $\halfcan$ is a half canonical divisor, 
\[
	2\halfcan=D+\sum \subhalf{e_i}P_i
\] 
where 
\[
	D\sim K_X+\Delta\sim -2\infty+ \Delta
\] is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

@@

\begin{example}
Here we demonstrate, in detail, that case (b) ($\sigma_2 =
(0; 3, 7, 7; 0), J_2 = \{3, 4\}$) satisfies the conditions.
We have that $K_{\sx'} \sim -(\infty) + \frac{1}{3} P_1 + \frac{3}{7}
P_2 + \frac{3}{7} P_3$ for some points $P_i \neq \infty$.
We exhibit a (minimal) presentation for $R'$.

Notice that $\deg \lfloor d K_{\sx'} = -d + \lfloor \frac{d}{3}
\rfloor + 2 \lfloor \frac{3d}{7} \rfloor$.

By Max Noether's theorem over $\BP^1$ (see Lemma 3.1.1 of Voight
and Zureick-Brown \cite{vzb:stacky}),

\[
	H^0 (\sx', \lfloor 21 K_{\sx'} \rfloor) \otimes H^0 (\sx', \lfloor
	(d - 21) K_{\sx'} \rfloor) \twoheadrightarrow = H^0 (\sx', \lfloor
	d K_{\sx'} \rfloor)
\]

\noindent
is surjective whenever $\deg (\lfloor (d - 21) K_{\sx'} \rfloor)
\geq 0$. Also, $\sigma_2$ has saturation $s = 5$ as noted in
Subsection ~\ref{ssec:g_0_saturation}. Thus, $R'$ is generated up to
degree $21 + (s - 1) = 24$.

For $d = 0, 1, 2, \ldots$ we have

\[
	dim H^0 (\sx', d K_{\sx'}) = 1, 0, 0, 1, 0, 1, 1, 2, 1, 1, 2, 1, 3, 2, 3, \ldots
\]

\noindent
so $R'$ must have some generators $x_{1, 3}, x_{1, 5}, x_{1, 7},
x_{1, 7}$. Let $(-ord_{P_1}(x_j), -ord_{P_2}(x_j), -ord_{P_3}(x_j))$
represent the pole orders of $x_i$ at each point. Then these generators
must have pole orders $(1, 1, 1), (1, 2, 2), (2, 3, 2), (2, 2, 3)$
respectively.

\end{example}

@@

\begin{longtable}
	{| c | c | c || c | c |}
	\hline
	Case & $\sigma$ & $J$ & $\deg \lfloor e_i L \rfloor$ & $\max_{d \geq 0} \#S_{(\sigma, J)}(i)$\\
	\hline
	\hline

@@

\section{Further Research}
\begin{itemize}
\item More on fractional modular forms
\item Make explicit generators and relations for base case of non-stacky higher genus half-canonical divisors
\end{itemize}

>>>>

\begin{prop}
\label{prop:sat_two_induction}
Let $(\sx, \Delta, L)$ and $(\sx', \Delta, L')$ be log stacky spin curves. 
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted log stacky curves so that $\sx$ has signature 
$\sigma = (g;e_1,\ldots, e_r;\delta)$ with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.
Let $R_{L'} = R' = k[x]/I'= k[x_2, x_3
, x_5,\ldots, x_m]/I'$ and let $L = L' + \frac{1}{
3}P.$ Suppose $L$ and $L'$ satisfy $sat(Eff(L')) = sat(Eff(L)) = 2.$
Suppose $\deg x_i = i$ for $i \in \{2, 3\}$ and that the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_3 \prec 
		y_4$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 3 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{align}
\label{eqn:sat_two_add_generator}
	\begin{split}
		&\langle y_3^ax_3^b x_4^\epsilon \mid a \geq 0, b 
		\geq 0, \epsilon \in \{0, 1\}\rangle \\
		&+\langle y_3^ay_4 \mid a \geq 0 \rangle,
	\end{split}
\end{align}

@@

\begin{align*}
	y_4x_j &- A_j y_3 w_j & \text{ so that } 3 \leq j \leq m \text{ and } \deg w_j = \deg x_j + 1\\
	y_4^2 &+ B y_3^2 x_2 \\
\end{align*}

@@

Let $\sx$ be a tame, seperably rooted stacky curve of genus 1. Let $P,Q$ are distinct hyperelliptic fixed points, and $R_1,\ldots, R_n$ be points on $\sx$. In this section, will compute the canonical ring $R_L$ in the cases that
\begin{enumerate}
	\item $L = 0$
	\item $L = P-Q$
	\item $L = P - Q + \frac{1}{3}R$
	\item $L = P - Q + \sum_{i=1}^{n}\frac{1}{3}R_i$
\end{enumerate}
In addition to being illustrative examples, these will serve as the base cases of induction, in order to compute generators and relations for an arbitrary genus 1, stacky log half canonical divisor. In each of the following examples, we will give explicit generators and relations, and in case $(3)$ and $(4)$ we check the corresponding signature is admissible.

\begin{example}
\label{eg:g_1_triv}
In the case $L = 0,$ we see immediately that $R_L \cong k[x]$ with $x$ a generator in degree 1.
\end{example}

\begin{example}
\label{eg:g_1_half_can}
In the case $L=P - Q.$ Then, we have $R_{(2k+1)L} \cong 0$ while $R_{2kL} \cong k$, with a single generator in degree 2. Therefore, $R_L \cong k[y]$, where $y$ is a generator in degree 2. 
\end{example}

\begin{example}
\label{eg:g_1_deg_third}
Let $L = P - Q + \frac{1}{3}R$. Since $P \neq Q$, we see $h^0(L) = 0,$ and we know $h^0(2L) = h^0(K) = 1$. For $n > 2$, we have $\deg nL > 0$ and so by Riemann--Roch, we have $h^0(nL) = \deg \lfloor nL/3 \rfloor.$ Next, 
We find that $R_L \cong k[u,x,z]/(z^2 - ux^4),$ where $\deg u = 2, \deg x = 3, \deg z = 7$ so that $\di u|_L = -2P + 2Q, \di x|_L = -3P + 3Q - R, \di z|_L = -7P + 7Q + 2R.$
\end{example}

\begin{example}
Let $L = P - Q + \frac{1}{3}R_1 + \frac{1}{3}R_2$. Here, we claim $R_L \cong k[u,x,y_3,y_4]/(xy_3- \alpha uy_4,y_4^2 - \beta x^2 u - \gamma y^2u)$ where $\alpha,\beta,\gamma \in k$, and $\deg u = 2, \deg x = \deg y_3 = 3, \deg y_4 = 4,$ where $u,x$ are as in Example ~\ref{eg:g_1_deg_third} and $y_3, y_4$ are as in Proposition ~\ref{prop:sat_two_induction}. \todo{Explain why this forms a system of generators and relations, and also state more clearly what $y_i$ are.}
\end{example}

\begin{example}
\label{}
Let $L = P - Q + \frac{e-1}{2e}R_1.$ Here, we claim $R_L \cong k[u,y_3, y_5, \ldots, y_e]/I_e,$ where we define $I_5 = y_3^4 - uy_5^2$ and $I_e$ is inductively defied via Theorem ~\ref{thm:ram_order_ind} so that $in(I_e) = in(I_{e-2})k[u, y_3,\ldots, y_e] + \langle y_e z \mid z \neq y_e, y_{e-2} \rangle.$ 
Furthermore, $\deg u = 2, \deg y_i = i$.
\todo{Note that here there is a quartic generator of $I$ in degree 12, providing exceptions to minimality by quadrics and an exception to the relation bound when $e = 5$.}
\end{example}

@@

@@

\todo{make sure to switch poles and zeros where appropriate}
\begin{lem}
\label{lem:semicanonical_generation}\todo{include that this was as shown in Neves, with degree 6 omitted}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K$. Then, for $n \geq 5$, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective.
In particular, $R_L$ is generated in degree at most $6$.
\end{lem}
\begin{proof}
Therefore, to show there are no new generators in degree at least $7$, it suffices to show that for $n \geq 5$, the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, let $s_1,s_2 \in H^0(2L) \cong H^0(K)$ be two independent sections. 
Since $K$ is basepoint free, so is the vector subspace $V = ks_1 \oplus k s_2 \subset H^0(K)$. 
Then, by the basepoint free pencil trick, we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$
We wish to show $f$ is surjective.
Note that $\dim \ker f = \dim H^0((n-2)L) = (n-3)(g-1)$ using Riemann--Roch and the assumption that $n \geq 5$.
Additionally, $\dim V \otimes H^0(nL) = 2 \cdot (n-1)(g-1)$, again using Riemann--Roch.
Therefore, $$\dim \im f = 2 \cdot (n-1)(g-1) -(n-3)(g-1) = (n+1)(g-1) = \dim H^0((n+2)L),$$
again using Riemann-Roch, implying $f$ is surjective.
\end{proof}

@@

\begin{cor}\label{cor:genus_0_posiive_delta}
Let $\halfcan$ be a stacky log spin canonical divisor of signature $(0,e_1,...,e_n,\delta)$ with $\delta>0$.  Then 
\end{cor}
\begin{proof}
Since $\halfcan$ is a half canonical divisor, 
\[
	2\halfcan=D+\sum \subhalf{i}P_i
\] 
where 
\[
	D\sim K_X+\Delta\sim -2\infty+ \Delta
\] is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

@@

\begin{example}
Here we demonstrate, in detail, that case (b) ($\sigma_2 =
(0; 3, 5, 9; 0), J_2 = \{3\}$) satisfies the conditions.
\end{example}

@@

\begin{longtable}
	{| c | c | c || c | c |}
	\hline
	Case & $\sigma$ & $J$ & $\deg \lfloor e_i L \rfloor$ & $\max_{d \geq 0} \#S_{(\sigma, J)}(i))$\\
	\hline
	\hline

@@
